Formalizing Stack Safety as a Security Property

### 2022-11-06 submission5/noexceptions.tex
fix minor latex issue
edits and comments

The set of special operations that are recognized in this way are termed
{\it security operations}. Our properties are phrased as a {\it security semantics},
which extends the underlying machine with additional context about its
security principals and which registers and memory they require to be secure.
This context evolves dynamically through the execution of security operations.
This notion of security may be phrased as
predicates on future states, e.g. assertions of the form
``When control returns to me...'', or relations on traces of future execution
(hyper-properties).

Unlike approaches that give a safe-by-construction semantics
for various security operations and then prove secure compilation to the underlying
enforcement mechanism, our machine is explicitly not safe unless the enforcement
is correctly instantiated; the security semantics gives us the means to judge whether
an enforcement scheme implements stack safety or not.

We will first give a detailed example of a security semantics for a simple setting
in which our security operations are restricted to making calls and returns, and allocating
private memory within the current function activation. We then extend this model
to the full system that we test, which features:
\begin{itemize}
\item Function calls and returns, with caller- and callee-saved registers
\item Allocation of strictly private and strictly public regions on the stack
\item Arguments spilled onto the stack and/or passed by reference
\item Exceptions
\item Tail-call Elimination
\end{itemize}

@@

We must trust that our method of distinguishing security operations is accurate; if it
involves labels placed on code by a compiler, that means trusting that the compiler placed
those labels correctly. If operations occur that are not recognized, those operations
might not be guaranteed to protect their principals---for instance, an unlabeled call
might not protect the caller's data from the callee. On the flip side, applying an incorrect
label most likely means that the property becomes too strong to be enforced.

@@

\begin{figure}
  \begin{tabular}{| l | l | l |}
    \hline
    Set / & Names & Purpose \\
    Class & & \\
    \hline
    \(\mathit{CLR}\) / & {\tt t0} -- {\tt t6} & Caller-saved temps \\
    \(\unsealed\) & {\tt a0} -- {\tt a1} & Caller-saved args / return vals \\
    & {\tt a2} -- {\tt a7} & Caller-saved args \\
    \hline
    \(\mathit{CLE}\) / & {\tt s0} -- {\tt s7} & Callee-saved \\
    \(\sealed\) & {\tt ra} & Return Address \\
    & {\tt sp} & Stack Pointer \\  
    \hline
    \(\mathit{PUBLIC}\) / & {\tt pc} & Program Counter \\
    \(\public\) & & \\
    \hline
  \end{tabular}
  \caption{RISC-V register set}
  \label{fig:RISCVregs}
\end{figure}

Collectively addresses and registers are {\em state elements} \(\component\)
in the set \(\COMPONENTS ::= \WORDS + \REGS\).
%
{\em Machine states} are drawn from a set \(\MACHS\) ranged over by \(\mach\)
that defines a mapping \(\mach[\component]\) from state elements to words.
{\em Events} are drawn from a set \(\OBSS\) and ranged over by \(\obs\), with a
special silent event \(\tau\).
The machine step function
\(\mach \xrightarrow{\overline{\psi},\obs} \mach' \in \MACHS \rightarrow
\MACHS \times \mathit{list} ~ \Psi \times \OBSS\)
is labeled by an ordered list of security operations, defined below,
and an event.
While the tag-based enforcement mechanism may cause the machine to
failstop, we model this as the machine silently diverging by remaining in the
same state. Since failstops are not distinguished from normal execution,
our properties will suffer the limitations typical of
{\it termination insensitive} systems \cite{}. TODO: more detailed explanation later.

\subsection{Security Operations, Principals, and Security State}

Security operations consist of calls (including tailcalls), returns,
allocations and deallocations. Calls and tailcalls identify the address of the call's target,
and the argument registers. Allocations and deallocations are phrased in terms of a pair
of an offset from the stack pointer and a size. Returns don't require any additional information.
%
\begin{align*}
  \psi \in \Psi ::= & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa} &
  sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
  | & \mathbf{tailcall} ~ \addr_{target}  ~ \overline{\reg_{args}} ~ \overline{sa} &
  sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
  | & \mathbf{return} & \\
  | & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
  | & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
\end{align*}
%
Security principals are function activations: at any given time, pending activations
have security requirements that are recorded in a call stack, and the current
activation tracks context information that will inform its requirements when
it makes a call. Each activation has a {\it view}
of the system that maps each state element to a {\it security class}:
\[sc \in SEC ::= \sealed | \unsealed | \object | \public\]
\[V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\]

@@

For convenience, we will abbreviate multiple updates of a mapping
\(V[\component_0 \mapsto sc_0][\component_1 \mapsto sc_1]\dots\)
as \(V \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
are fomulae with free variables bound to sets in \(C\).

We represent the call stack with a list of views representing pending inactive
principals, ranged over by \(\sigma\). A {\it context} is a pair of the current
activation's view and that list. The initial context is \(\context_0 = (V_0, \emplist)\).
\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ VIEW}\]

Every security operation manipulates the call  stack via a function
\(Op : \MACHS \rightarrow \CONTEXTS \rightarrow \psi \rightarrow \CONTEXTS\).

@@

In addition to its elements, a system may contain additional ``policy'' state that carries
information relevant to a hardware enforcement mechanism. We assume that policy state
does not influence execution, except to cause the system to failstop, which is modeled
as a state stepping to itself. A failstop is therefore indistinguishable from silent
divergence.

Formally, we can treat our machine state as
\(\mach = (\mach[\cdot],\langle) \mach \rangle \in \MACHS
::= (\COMPONENTS \rightarrow \WORDS) \times \POLS\) for some \(\POLS\).
Then \(\langle \mach \rangle\) is the security state. We assume that if
\(\mach[\component] = \nach[\component]\) for all \(\component\),
\(\mach \stepstounder{\bar{\psi_1}} \mach'\), and \(\nach \stepstounder{\bar{\psi_2}} \nach'\),
then \(\bar{\psi_1} = \bar{\psi_2}\) and either
\(\mach = \mach'\), \(\nach = \nach'\), or \(\mach'[\component] = \nach'[\component]\)
for all \(\component\).

@@

In this section we consider a simple setting with calls, returns, and private allocations.
Figure \ref{fig:main} gives a C function {\tt main}, and its compiled RISC-V code,
that takes a secret as its argument and initializes a variable to contain
potentially sensitive data. We will assume for this system that events are stores
to a special address, called {\tt out}.

{\tt main} then calls another function {\tt f},
and afterward may decide to output its secret based on the contents of
its sensitive data. In this case, since {\tt sensitive} is initialized to 0,
it should not---instead it will publish the results of {\tt f}.

\begin{figure}
  \begin{subfigure}{\columnwidth}
    {\tt
      void main(int secret) \{

@@

      ~ ~ if (sensitive == 42) \{

      ~ ~ ~ ~ *out = secret;

      ~ ~ \} else \{

      ~ ~ ~ ~ *out = res;

      ~ ~ \}

      \}}
  \end{subfigure}
  \begin{subfigure}{\columnwidth}
    \begin{tabular}{r l | l}
      \labeledrow{0:}{addi sp,sp,-20}{\(\mathbf{alloc} ~ (-20,20)\)}
      \labeledrow{4:}{sd ra,12(sp)}{}
      \labeledrow{8:}{sw a1,8(sp)}{}
      \labeledrow{12:}{sw zero,4(sp)}{}
      \labeledrow{16:}{jal f,ra}{\(\mathbf{call} ~ \mathtt{f} ~ \emplist ~ \emplist\)}
      \labeledrow{20:}{sw r0,0(sp)}{}
      \labeledrow{24:}{lw a4,4(sp)}{}
      \labeledrow{28:}{li a5,42}{}
      \labeledrow{32:}{bne a4,a5,L1}{}
      \labeledrow{36:}{lw a0,8(sp)}{}
      \labeledrow{40:}{sw a0,out}{}
      \labeledrow{44:}{j L2:}{}
      \labeledrow{L1, 48:}{lw r0,0(sp)}{}
      \labeledrow{52:}{sw a0,out}{}
      \labeledrow{L2, 56:}{ld ra,12(sp)}{}
      \labeledrow{60:}{addi sp,sp,20}{\(\mathbf{dealloc} ~ (0,20)\)}
      \labeledrow{64:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
  \end{subfigure}
  \begin{subfigure}{\columnwidth}
    \includegraphics[width=\columnwidth]{stacklayout.png}
  \end{subfigure}

@@

Suppose that {\tt f} is not a (valid) C function at all, but an attacker seeking
to leak {\tt secret}! It might do so in a number of ways, shown as snippets of
assembly code in Figure \ref{fig:f}.

In Figure \ref{subfig:direct}, {\tt f} takes an offset from the stack
pointer, accesses {\tt secret}, and directly outputs it. But more
subtly, even if somehow prevented from outputting {\tt secret} directly, {\tt f}
can instead return that value so that {\tt main} stores it to {\tt out},
as in Figure \ref{subfig:indirect}.

Beyond simply reading {\tt secret}, the attacker might overwrite {\tt sensitive}
with 42, guaranteeing that {\tt main} publishes its own secret unintentionally
(Figure \ref{subfig:integrity})!
Attacks of this kind do not violate {\tt main}'s confidentiality, but its
{\it integrity}.
And in Figure \ref{subfig:WBCF}, the attacker attempts to return to the
wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
violating the program's {\it well-bracketed control flow} (WBCF.)

\begin{figure}
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{lw a4,8(sp)}{}
      \labeledrow{104:}{sw a4,out}{}
      \labeledrow{108:}{li a0,1}{}
      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \caption{Leaking {\tt secret} directly}
    \label{subfig:direct}
  \end{subfigure}  
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{lw a4,8(sp)}{}
      \labeledrow{104:}{mov a4,a0}{}
      \labeledrow{108:}{sw zero,-4(sp)}{}
      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \caption{Leaking {\tt secret} indirectly}
    \label{subfig:indirect}
  \end{subfigure}  
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{li a5,42}{}
      \labeledrow{104:}{sw a5,4(sp)}{}
      \labeledrow{108:}{li a0,1}{}
      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \subcaption{Attacking {\tt sensitive}}
    \label{subfig:integrity}
  \end{subfigure}
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{addi ra,ra,20}{}
      \labeledrow{104:}{nop}{}
      \labeledrow{108:}{nop}{}
      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \subcaption{Attacking control flow}
    \label{subfig:WBCF}
  \end{subfigure}

@@

Figure \ref{fig:exec1} shows how the security context
develops over several steps, with labels on the steps that allocate space and then make
a call. Both program counter and stack pointer are initially 0, and the stack pointer
is shown for convenience.
\(\xrightarrow{\overline{\psi}}\) transitions are written \(\downarrow \overline{\psi}\)
in this diagram; events are omitted.

The first step  allocates words for {\tt secret}, {\tt sensitive}, and {\tt res}, as well
as two words for the return address. This will have the
effect of marking those bytes \(\object\), assuming they were previously
\(\unsealed\).

\begin{figure*}
  \begin{subfigure}[t]{.5\textwidth}
    \vskip 0pt
    \begin{tabular}{| l | l || l |}
      \hline
      Step & \(\SP\) & \(\context\) \\
      \hline
      0 & 0 & \(V_0, \emplist\) \\
      \hline
      \multicolumn{3}{l}{\(\downarrow [\mathbf{alloc} ~ (-20,20)]\)} \\
      \hline
      1 & -20 & \(V_1 = V_0 \llbracket -20 \ldots -1, \mapsto \object \rrbracket, \emplist\) \\
      \hline
      \multicolumn{3}{l}{\(\vdots\)} \\
      \hline
      4 & -20 & \(V_1, \emplist\) \\
      \hline
      \multicolumn{3}{l}{\(\downarrow [\mathbf{call} ~ 100 ~ \emplist ~ \emplist]\)} \\
      \hline
      5 & -20 & \(V_2 = V_1 [\mathtt{a0} \mapsto \unsealed],[V_1]\) \\
      \hline
      \multicolumn{3}{l}{\(\vdots\)} \\
      \hline
      8 & -20 & \(V_2,[V_1]\) \\
      \hline
      \multicolumn{3}{l}{\(\downarrow [\mathbf{return}]\)} \\
      \hline
      9 & -20 & \(V_1, \emplist\) \\
      \hline
    \end{tabular}
  \end{subfigure}
  \begin{subfigure}[t]{.4\textwidth}
    \vskip 0px
    \includegraphics[width=\columnwidth]{stack1.png}
    \includegraphics[width=\columnwidth]{stack2.png}
    \includegraphics[width=\columnwidth]{stack3.png}
    \includegraphics[width=\columnwidth]{stack4.png}
  \end{subfigure}

\caption{Execution through return from {\tt f}}
\label{fig:exec1}
\end{figure*}

On a call, the formerly active principal's record is pushed onto the inactive list.
Its return target is the return address of the call, 
and the stack pointer target is the stack pointer at the moment of call.
The callee's view is updated from the caller's such that all \(\object\) locations
become \(\sealed\). (In more complex settings, objects that are intentionally passed
to the callee will not get sealed, but for now we assume no sharing of memory between activations.)
For registers, \(\overline{\reg_{args}}\) tells us which registers are used as arguments,
in this case {\tt a0}. These are mapped to \(\public\), while any non-argument, caller-saved
registers are mapped to \(\unsealed\). All callee-save registers are always \(\sealed\) and
the program counter \(\public\), as they were initialized.

On step 8, {\tt f} returns, and we restore the topmost inactive principal,
{\tt main}'s principal. Now we will discuss the intuition behind how this
structure translates into stack safety properties that rule out different
classes of bad behavior.

@@

But what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or
\(\SP \neq -20\)), as would be the case in attack (d)? We consider this to
violate well-bracketed control flow (WBCF). WBCF is a relationship between
call steps and their corresponding return steps: after the return, the program
counter should be at the instruction following the state before the call,
and the stack pointer should be the same as it was before the call.

Consider the attack in Figure \ref{subfig:WBCF}: the attacker adds
20 to the return address and then returns, thus bypassing the check and outputting
{\tt secret}. Before the call, the program counter is 16 and the stack pointer is 20,
so we define a predicate on states \(ret\) that holds when the
program counter and stack pointer are both 20.

Then we identify the state immediately following the return, if one occurs ---
this will be the first state in which the pending call stack is smaller than it was
inside of the call.
WBCF requires that if \(s\) after that return, \(ret ~ s\) will hold.
For nested calls, where the pending stack is initial larger, the same principal
applies: \(ret ~ s\) must hold the next time the pending stack is the same size or smaller.

To demonstrate why WBCF also cares about the stack pointer, consider an attack which
returns with the stack pointer at -12 instead of -20. Now, compare the layout of the
stack relative to the stack pointer as it should be (\(\SP\)) to the one that results by
increasing the stack pointer before return (\(\SP'\)):

\begin{tabular}{| l | l | l | l | l |}
  \multicolumn{1}{r}{\(\SP \downarrow\)} &
  \multicolumn{2}{r}{\(\SP' \downarrow\)} \\
  \hline
  res & sens & sec & ra & ra \\
  \hline
\end{tabular}

@@

In this scenario, when attempting to read {\tt sensitive}, {\tt main} will
read part of the return address, and then it will attempt to output
{\tt res}, but instead output {\tt secret}! Even absent other
kinds of data protection, the stack pointer {\it must} be restored
for the program to behave predictably.

>>>>

The set of special operations that are recognized in this way are termed
{\it security-relevant operations}. Our properties are phrased as a {\it security semantics},
which extends the underlying machine with additional context about its
security principals and which registers and memory they require to be secure.
This context evolves dynamically through the execution of security-relevant operations.
This notion of security may be phrased as
predicates on future states, e.g. assertions of the form
``When control returns to me...'', or relations on traces of future execution
(hyper-properties).

Unlike approaches that give a safe-by-construction semantics
for various security-relevant operations and then prove secure compilation to the underlying
enforcement mechanism, our machine is explicitly not safe unless the enforcement
is correctly instantiated; the security semantics gives us the means to judge whether
an enforcement scheme implements stack safety or not.

We will first give a detailed example of a security semantics for a simple setting
in which our security-relevant operations are restricted to making calls and returns, and allocating
private memory within the current function activation. We then extend this model
to the full system that we test, which features:
\begin{itemize}
\item Function calls and returns, with caller- and callee-saved registers
\item Allocation of strictly private and strictly public regions on the stack
\item Arguments spilled onto the stack and/or passed by reference
\item Exceptions
\item Tail-call Elimination
\end{itemize}

@@

We must trust that our method of distinguishing security-relevant operations is accurate; if it
involves labels placed on code by a compiler, that means trusting that the compiler placed
those labels correctly. If operations occur that are not recognized, those operations
might not be guaranteed to protect their principals---for instance, an unlabeled call
might not protect the caller's data from the callee. On the flip side, applying an incorrect
label most likely means that the property becomes too strong to be enforced.

@@

\begin{figure}
  \begin{tabular}{| l | l | l |}
    \hline
    Set / & Names & Purpose \\
    Class & & \\
    \hline
    \(\mathit{CLR}\) / & {\tt t0} -- {\tt t6} & Caller-saved temps \\
    \(\unsealed\) & {\tt a0} -- {\tt a1} & Caller-saved args / return vals \\
    & {\tt a2} -- {\tt a7} & Caller-saved args \\
    & {\tt ra} & Return Address \\
    \hline
    \(\mathit{CLE}\) / & {\tt s0} -- {\tt s11} & Callee-saved \\
    \(\sealed\) & {\tt sp} & Stack Pointer \\  
    \hline
    \(\mathit{PUBLIC}\) / & {\tt gp} & Global Pointer  \\
    \(\public\) & {\tt tp} & Thread Pointer \\
    \hline
  \end{tabular}
  \caption{RISC-V integer register set}
  \label{fig:RISCVregs}
\end{figure}

Collectively addresses and registers are {\em state elements} \(\component\)
in the set \(\COMPONENTS ::= \WORDS + \REGS\).
%
{\em Machine states} are drawn from a set \(\MACHS\) ranged over by \(\mach\)
that defines a mapping \(\mach[\component]\) from state elements to words.
{\em Events} are drawn from a set \(\OBSS\) and ranged over by \(\obs\), with a
special silent event \(\tau\).
The machine step function
\(\mach \xrightarrow{\overline{\psi},\obs} \mach' \in \MACHS \rightarrow
\MACHS \times \mathit{list} ~ \Psi \times \OBSS\)
is labeled by an ordered list of security-relevant operations, defined below,
and an event.
While the tag-based enforcement mechanism may cause the machine to
failstop, we model this as the machine silently diverging by remaining in the
same state. Since failstops are not distinguished from normal execution,
our properties will suffer the limitations typical of
{\it termination insensitive} systems \cite{}. TODO: more detailed explanation later.

\subsection{Security-Relevant Operations, Principals, and Security State}

\apt{Maybe should drop the math displays from this section, and keep to words.}
Security-relevant operations consist of calls (including tailcalls), returns,
allocations and deallocations. \apt{No tailscalls for now?} Calls and tailcalls identify the address of the call's target,
and the argument registers. \apt{Either explain $\overline{sa}$ or omit it.} Allocations and deallocations are phrased in terms of a pair
of an offset from the stack pointer and a size. Returns don't require any additional information.
%
\begin{align*}
  \psi \in \Psi ::= & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa} &
  sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
  | & \mathbf{tailcall} ~ \addr_{target}  ~ \overline{\reg_{args}} ~ \overline{sa} &
  sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
  | & \mathbf{return} & \\
  | & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
  | & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
\end{align*}
%
Security principals are function activations: at any given time, pending activations
have security requirements that are recorded in a call stack, and the current
activation tracks context information that will inform its requirements when
it makes a call. Each activation has a {\it view}
of the system that maps each state element to a {\it security class}:
\[sc \in SEC ::= \sealed | \unsealed | \object | \public\]
\[V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\]

@@

For convenience, we will abbreviate multiple updates of a mapping
\(V[\component_0 \mapsto sc_0][\component_1 \mapsto sc_1]\dots\)
as \(V \llbracket A \mapsto B | C \rrbracket\), where \(A\) and \(B\)
are fomulae with free variables bound to sets in \(C\).
\apt{can omit this para}

A {\it context} is a pair of the current activation's view and 
a list of views representing the call stack (pending inactive
principals), ranged over by \(\sigma\).  
\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ VIEW}\]
The initial context is \(\context_0 = (V_0, \emplist)\).

\apt{Maybe don't need to describe $Op$ or $\stepstounder{\overline{\psi}}$ here, 
  and maybe not even combined states, although they are used at the moment.}
  Every security-relevant operation manipulates the call stack via a function
\(Op : \MACHS \rightarrow \CONTEXTS \rightarrow \psi \rightarrow \CONTEXTS\).

@@

\apt{As we discussed at length, it would certainly be nice to avoid separate policy state if possible.}
In addition to its elements, a system may contain additional ``policy'' state that carries
information relevant to a hardware enforcement mechanism. We assume that policy state
cannot influence execution, except possibly to cause the system to failstop, which is modeled
as a state stepping to itself. A failstop is therefore indistinguishable from silent
divergence.

Formally, we can treat our machine state as
\(\mach = (\mach[\cdot],\langle \mach \rangle) \in \MACHS
::= (\COMPONENTS \rightarrow \WORDS) \times \POLS\) for some \(\POLS\).
\apt{This overloaded notation doesn't work. $\mach$ can be either a pair or a map, not both.
Note that earlier text strongly implies that it is a map.}
Then \(\langle \mach \rangle\) is the security state. We assume that if
\(\mach[\component] = \nach[\component]\) for all \(\component\),
\(\mach \stepstounder{\bar{\psi_1}} \mach'\), and \(\nach \stepstounder{\bar{\psi_2}} \nach'\),
then \(\bar{\psi_1} = \bar{\psi_2}\) and either
\(\mach = \mach'\), \(\nach = \nach'\), or \(\mach'[\component] = \nach'[\component]\)
for all \(\component\).

@@

In this section we consider a simple setting with calls, returns, and private allocations.
Figure \ref{fig:main} gives the C code and corresponding compiled RISC-V code
for a function {\tt main}. (Code is compiled using the standard RISC-V
calling conventions~\cite{??}; the stack grows towards lower addresses. We assume that
{\tt main} begins at address 0 and its callee {\tt f} at address 100.)

{\tt main} takes a secret as its argument and initializes a variable to contain
potentially sensitive data. We assume that observable events are stores
to a special address, called {\tt out}. 
{\tt main} calls another function {\tt f},
and afterward may decide to output its secret based on the contents of
its sensitive data. In this case, since {\tt sensitive} is initialized to 0,
it should not---instead it will publish the results of {\tt f}.

\begin{figure}
  \begin{subfigure}{\columnwidth}
    {\tt
      volatile int out;

      void main(int secret) \{

@@

      ~ ~ if (sensitive == 42)

      ~ ~ ~ ~ out = secret;

      ~ ~ else

      ~ ~ ~ ~ out = res;

      \}}
  \end{subfigure}
  \begin{subfigure}{\columnwidth}
    \begin{tabular}{r l | l}
      \labeledrow{0:}{addi sp,sp,-20}{\(\mathbf{alloc} ~ (-20,20)\)}
      \labeledrow{4:}{sd ra,12(sp)}{}
      \labeledrow{8:}{sw a1,8(sp)}{}
      \labeledrow{12:}{sw zero,4(sp)}{}
      \labeledrow{16:}{jal f,ra}{\(\mathbf{call} ~ \mathtt{f} ~ \emplist ~ \emplist\)}
      \labeledrow{20:}{sw a0,0(sp)}{}
      \labeledrow{24:}{lw a4,4(sp)}{}
      \labeledrow{28:}{li a5,42}{}
      \labeledrow{32:}{bne a4,a5,L1}{}
      \labeledrow{36:}{lw a0,8(sp)}{}
      \labeledrow{40:}{sw a0,out}{}
      \labeledrow{44:}{j L2:}{}
      \labeledrow{L1, 48:}{lw a0,0(sp)}{}
      \labeledrow{52:}{sw a0,out}{}
      \labeledrow{L2, 56:}{ld ra,12(sp)}{}
      \labeledrow{60:}{addi sp,sp,20}{\(\mathbf{dealloc} ~ (0,20)\)}
      \labeledrow{64:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
  \end{subfigure}
  \begin{subfigure}{\columnwidth}
    \includegraphics[width=\columnwidth]{stacklayout.png}
  \end{subfigure}

@@

Suppose that {\tt f} is not a (valid) C function at all, but an attacker seeking
to leak {\tt secret}! It might do so in a number of ways, shown as snippets of
assembly code in Figure \ref{fig:f}.\footnote{We arrange that all the snippets
have the same length just to simplify some later explanations.}
%
In Figure \ref{subfig:direct}, {\tt f} takes an offset from the stack
pointer, accesses {\tt secret}, and directly outputs it. But more
subtly, even if somehow prevented from outputting {\tt secret} directly, {\tt f}
can instead return that value so that {\tt main} stores it to {\tt out},
as in Figure \ref{subfig:indirect}.
%
Beyond simply reading {\tt secret}, the attacker might overwrite {\tt sensitive}
with 42, guaranteeing that {\tt main} publishes its own secret unintentionally
(Figure \ref{subfig:integrity})!
Attacks of this kind do not violate {\tt main}'s confidentiality, but its
{\it integrity}.
And in Figure \ref{subfig:WBCF}, the attacker arranges to return to the
wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
violating the program's {\it well-bracketed control flow} (WBCF.)

\begin{figure}
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{lw a4,8(sp)}{}
      \labeledrow{104:}{sw a4,out}{}
      \labeledrow{108:}{li a0,1}{}
      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \caption{Leaking {\tt secret} directly}
    \label{subfig:direct}
  \end{subfigure}  
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{lw a4,8(sp)}{}
      \labeledrow{104:}{mov a4,a0}{}
      \labeledrow{108:}{sw zero,-4(sp)}{}
      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \caption{Leaking {\tt secret} indirectly}
    \label{subfig:indirect}
  \end{subfigure}  
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{li a5,42}{}
      \labeledrow{104:}{sw a5,4(sp)}{}
      \labeledrow{108:}{li a0,1}{}
      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \subcaption{Attacking {\tt sensitive}}
    \label{subfig:integrity}
  \end{subfigure}
  \begin{subfigure}[b]{\columnwidth}
    \vspace{\abovedisplayskip}
    \begin{tabular}{r l | l}
      \labeledrow{100:}{addi ra,ra,16}{}
      \labeledrow{104:}{nop}{}
      \labeledrow{108:}{nop}{}
      \labeledrow{112:}{jalr ra}{\(\mathbf{return}\)}
    \end{tabular}
    \subcaption{Attacking control flow}
    \label{subfig:WBCF}
  \end{subfigure}

@@

The security semantics for this program is based
on the security-relevant events noted in the right columns of Figures~\ref{fig:main}
and~\ref{fig:f}, namely execution of instructions that allocate or deallocate space,
make a call, or make a return.
Figure \ref{fig:exec1} shows how the corresponding security context
develops over the first few steps of the program. We assume execution begins
at the start of {\tt main} (\(\PCname = 0\)) and that the stack pointer is initially at
address 1000. %  the \(PC\) and \(SP\) at each step are shown for clarity.
\(\xrightarrow{\overline{\psi}}\) transitions are numbered and
written \(\downarrow \overline{\psi}\) in this diagram; events are omitted.

Step 1 allocates a word each for {\tt secret}, {\tt sensitive}, and {\tt res}, as well
as two words for the return address. This will have the
effect of marking those bytes \(\object\), assuming they were previously
\(\unsealed\). (We use \(V\llbracket\cdot\rrbracket\) to denote updates to \(V\).\apt{To avoid confusion with list brackets.} )
%
\begin{figure*}
  \begin{subfigure}[t]{.55\textwidth}
    \vskip 0pt
    \begin{tabular}{|r|r||l|}
      \hline
      \(\PCname\) & \(\SP\) & \(\context\) \\
      \hline
      0 & 1000 & \(V_0, \emplist\) \\
      \hline
      \multicolumn{3}{l}{\(1 \downarrow [\mathbf{alloc} ~ (-20,20)]\)} \\
      \hline
      4 & 980 & \(V_1 = V_0 \llbracket 980..999 \mapsto \object\rrbracket, \emplist\) \\
      \hline
      \multicolumn{3}{l}{\(2 \downarrow []\)} \\
      \multicolumn{3}{l}{\(\ \ \vdots\)} \\
      \multicolumn{3}{l}{\(4 \downarrow []\)} \\
      \hline
      16 & 980 & \(V_1, \emplist\) \\
      \hline
      \multicolumn{3}{l}{\(5 \downarrow [\mathbf{call} ~ 100 ~ \emplist ~ \emplist]\)} \\
      \hline
      100 & 980 & \(V_2 = V_1 \llbracket 980..999 \mapsto \sealed, \mathtt{a0} \mapsto \unsealed\rrbracket,[V_1]\) \\
      \hline
      \multicolumn{3}{l}{\(6 \downarrow []\)} \\
      \multicolumn{3}{l}{\(\ \ \vdots\)} \\
      \multicolumn{3}{l}{\(8 \downarrow []\)} \\
      \hline
      112 & 980 & \(V_2,[V_1]\) \\
      \hline
      \multicolumn{3}{l}{\(9 \downarrow [\mathbf{return}]\)} \\
      \hline
      20(?) & 980  & \(V_1, \emplist\) \\
      \hline
    \end{tabular}
  \end{subfigure}
  \begin{subfigure}[t]{.4\textwidth}
    \vskip 0px
    \includegraphics[width=\columnwidth]{stack1.png}
    \includegraphics[width=\columnwidth]{stack2.png}
    \includegraphics[width=\columnwidth]{stack3.png}
    \includegraphics[width=\columnwidth]{stack4.png}
  \end{subfigure}

\caption{Execution up through the return from {\tt f}\apt{Label rows on right. Also show registers?}}
\label{fig:exec1}
\end{figure*}
%
At step 5, the formerly active principal's record is pushed onto the inactive list.
Its return target is the return address of the call, 
and the stack pointer target is the stack pointer at the moment of call.
The callee's view is updated from the caller's such that all \(\object\) locations
become \(\sealed\). (In more complex settings, objects that are intentionally passed
to the callee will not get sealed, but for now we assume no sharing of memory between activations.)
For registers, \(\overline{\reg_{args}}\) tells us which registers are used as arguments,
in this case just {\tt a0}. These are mapped to \(\public\), while any non-argument, caller-saved
registers remain \apt{right?} mapped to \(\unsealed\). All callee-save registers are always \(\sealed\)
as they were initialized\apt{???}.
At step 9, {\tt f} returns, and the topmost inactive principal, for {\tt main}, is restored.

We now show how this security semantiy can be used define notions of confidentiality,
integrity, and correct control flow in such a way that many classes of
bad behavior, including the attacks in Figure~\ref{fig:f}, are 
are detected as security violations.

@@

To begin with, what if {\tt f} returns to an unexpected place (i.e. \(\PCname \neq 20\) or
\(\SP \neq 980\)). We consider this to
violate well-bracketed control flow (WBCF). WBCF is a relationship between
call steps and their corresponding return steps: just after the return, the program
counter should be at the next instruction following the call,
and the stack pointer should be the same as it was before the call.

Consider the attack in Figure \ref{subfig:WBCF}: the attacker adds
16 to the return address and then returns, thus bypassing the {\tt if}-test in the code and outputting
{\tt secret}. Before the call, the program counter is 16 and the stack pointer is 980,
so we define a predicate \(ret\) \apt{suggest capitalizing these predicates}
on states that should hold just after the return, namely, the program counter
should be 20 and the stack pointer should be 980. \apt{Maybe write this symbolically.}

We can identify the point just after the return (if a return occurs)
as the first state in which the pending call stack is smaller than it was
just after the call.
WBCF requires that if \(s\) is the state at that point, then \(ret ~ s\) holds.
For nested calls, where the pending stack is initially larger, the same principle
applies: \(ret ~ s\) must hold the next time the pending stack is the same size or smaller.

To see why WBCF also cares about the stack pointer, consider an attack \apt{should this
  be shown as a fifth attack in the figure?} which
returns with \(\SP' = 998\) instead of the correct \(\SP = 980\), and recall
the layout of {\tt main}'s frame: 
\begin{tabular}{| l | l | l | l | l |}
  \multicolumn{1}{r}{\(\SP \downarrow\)} &
  \multicolumn{2}{r}{\(\SP' \downarrow\)} \\
  \hline
  res & sens & sec & ra & ra \\
  \hline
\end{tabular}

@@

\noindent
In this scenario, when attempting to read {\tt sensitive}, {\tt main} will
read part of the return address, and then it will attempt to output
{\tt res}, but instead output {\tt secret}! Even absent other
kinds of data protection, the stack pointer {\it must} be restored
for the program to behave predictably.
